### 2017-08-31 paper.tex
:hammer: Fix typos and figure labels

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{cleveref}
\usepackage{dsfont}
\RequirePackage{amsmath,amsfonts,amssymb}
\RequirePackage{booktabs}

@@

\Abstract{This paper investigates the effect of permeability on the runout characteristics of collapse of  granular columns on slopes in fluid. Two-dimensional sub-grain scale numerical simulations are performed to understand the flow dynamics of granular collapse in fluid. The Discrete Element (DEM) technique is coupled with the Lattice Boltzmann Method (LBM), for fluid-grain interactions, to understand the evolution of submerged granular flows. The fluid phase is simulated using Multiple-Relaxation-Time LBM (LBM-MRT) for numerical stability. In order to simulate interconnected pore space in 2D, a reduction in the radius of the grains (hydrodynamic radius) is assumed during LBM computations. The collapse of granular column in fluid is compared with the dry cases to understand the effect of fluid on the runout behaviour. A parametric analysis is performed to assess the influence of the granular characteristics (initial packing) on the evolution of flow and run-out distances for slope angle of \SI{5}{\degree}. In order to understand the effect of permeability on granular flow down a slope angle of \SI{5}{\degree}, the collapse of a granular column with an initial aspect ratio of 0.8 is simulated with different permeabilities. The hydrodynamic radius of a loosely packed granular column is varied from r = 0.7 R (high permeability), 0.75 R, 0.8 R, 0.85 R to 0.9 R (low permeability). The granular flow dynamics is investigated by analysing the effect of hydroplaning, water entrainment and viscous drag on the granular mass. The mechanism of energy dissipation, shape of the flow front, water entrainment and evolution of packing density is used to explain the difference in the flow characteristics of granular column collapse in fluid.}

@@

\section{Granular column collapse on slopes in fluid}
\subsection{Problem definition}
In this study, a 2D polydisperse system ($d_{max}/d_{min} = 1.8$) of circular discs in fluid was used to understand the behaviour of granular flows on inclined planes. As shown in~\cref{fig:setup}, an inclined gravity direction was varied to examine the slope angle effect. The soil column was modelled using 1000 discs of density \SI{2650}{\kg\per\cubic\meter} and a contact friction angle of \SI{26}{\degree}. The aspect ratio `a' is defined as the ratio of the initial height ($H_i$) to the width ($L_i$) of the column. A granular column of aspect ratio `a' of 0.8 was used. The collapse of the column was simulated inside a fluid with a density of \SI{1000}{\kg\per\cubic\meter} and a kinematic viscosity of \SI{1e-6}{\square\meter\per\second}. The choice of a 2D geometry has the advantage of cheaper computational effort than a 3D case, making it feasible to simulate very large systems. To model 3D permeability nature of spheres as 2D discs, a reduction in radius: a  hydrodynamic radius r = 0.9R was adopted only for LBM computations, as described earlier. Dry column collapse was also performed to study the effect of hydrodynamic forces on the runout distance. The runout distance is normalised with respect to the initial width of the column. The granular column collapse was allowed to flow down a slopes of \SI{5}{\degree}.

\begin{figure}[htpb]
\includegraphics[width=0.97\columnwidth]{figs/geometry}
\caption{Underwater granular collapse set-up}
\label{fig:setup}
\end{figure}

@@

\begin{figure}[htpb]
\includegraphics[width=0.97\columnwidth]{figs/dense_s5_r09}
\caption{Evolution of a dense sand column (aspect ratio 0.8) on a slope of \SI{5}{\degree}. }
\label{fig:dense-s5-r09}
\end{figure}

@@

\section{Effect of initial density - Loose versus Dense}
\citet{Topin2011} observed from CFD-DEM simulations of granular collapse on a horizontal plane in different viscous fluids that, for a given initial geometry, the run-out distance in the dry case is significantly higher than the submerged case, an observation similar to the experimental results of~\citet{Cassar2005}. For a given geometry, the initial volume fraction of the granular mass has a significant effect on the morphology of the granular deposits in fluid~\citep{Rondon2011a, Pailha2008}. In the dry case, inertia is responsible for the enhanced mobility at steeper slopes. In the submerged cases, however, the grain-inertial effects remain negligible because of the predominance of viscous effects on the granular dynamics~\citep{Topin2011}. This could explain the importance of the initial volume fraction on the runout behaviour of submerged granular columns. In this study, we examine the effect of initial packing density on the runout behaviour at \SI{5}{\degree} slope angle in both dry and submerged cases.

@@

\begin{figure}
\makebox[\linewidth][c]{
\begin{subfigure}[b]{0.95\textwidth}
	\centering
    \includegraphics[width=0.95\textwidth]{figs/loose_slope/LBM_660_Slope5_r07}
    \caption{r = 0.7 R}
    \label{fig:LBM_660_Slope5_r07}
\end{subfigure}
}\\
\makebox[\linewidth][c]{
\begin{subfigure}[b]{0.95\textwidth}
	\centering
    \includegraphics[width=0.95\textwidth]{figs/loose_slope/LBM_660_Slope5_r075}
    \caption{r = 0.75 R}
    \label{fig:LBM_660_Slope5_r075}
\end{subfigure}
} \\
\makebox[\linewidth][c]{
\begin{subfigure}[b]{0.95\textwidth}
	\centering
    \includegraphics[width=0.95\textwidth]{figs/loose_slope/LBM_660_Slope5_r08}
    \caption{r = 0.8 R}
    \label{fig:LBM_660_Slope5_r08}
\end{subfigure}
}\\
\makebox[\linewidth][c]{
\begin{subfigure}[b]{0.95\textwidth}
	\centering
    \includegraphics[width=0.95\textwidth]{figs/loose_slope/LBM_660_Slope5_r085}
    \caption{r = 0.85 R}
    \label{fig:LBM_660_Slope5_r085}
\end{subfigure}
}\\
\makebox[\linewidth][c]{
\begin{subfigure}[b]{0.95\textwidth}
	\centering
    \includegraphics[width=0.95\textwidth]{figs/loose_slope/LBM_660_Slope5_r09}
    \caption{r = 0.9 R}
    \label{fig:LBM_660_Slope5_r09}
\end{subfigure}
}
\caption{Evolution of the flow front at $t = 3\tau_c$ for different 
permeabilities (loose slope 5\si{\degree}).}
\label{fig:slope_loose_5}
\end{figure}

@@

\subsubsection{Water entrainment}
Comparing the snapshots of the granular flow at time $t=3\tau_c$ for the dense column cases~\cref{fig:dense-s5-r09} and the loose column cases~\cref{fig:slope_loose_5}, it can be seen that the overall shapes of the flow front are different.~\Cref{fig:slope-outline} shows the boundary of the flowing mass and its evolution with time for the slope angle \si{5}{\degree} case. It shows that the dense granular columns exhibit a more acute angle of attack (flow front) in contrast to a more parabolic profile for the loose granular columns. This difference in angle of flow front changes the ratio of amount of water entering the flowing granular mass to the water interacting with the surface of the granular mass.~\Cref{fig:water_entrainment_loose_dense} shows the horizontal velocity of the fluid and streamlines at the flow front of the dense and loose granular columns for the slope angle of \si{5}{\degree}. The density of streamlines entering the granular front is significantly higher for the loose column case, in comparison to the dense column case. This clearly shows the effect of the shape of flow front on water entrainment. More water entrainment results in increase in pore space (or decrease in solid fraction) at the flow front, which in turn reduces the interparticle shear resistance.

@@

\subsubsection{Turbulence and drag}
Closer observation of the surface of the flow front in the ~\cref{fig:dense-s5-r09} and the loose column cases~\cref{fig:slope_loose_5} cases shows that the number of turbulent vortices formed on the surface of the flow is significantly higher in the dense column cases than in the loose column cases. This can be attributed to the triangular profiles of the flowing mass of the dense column cases, in contrast to the more parabolic profiles observed in the loose granular column cases. This increase in the interaction between the flowing granular mass and the neighbouring fluid results in an increased momentum transfer and thus formation of turbulent vortices.~\Cref{fig:water_entrainment_a08_dense} shows higher density of streamlines at the top surface for the dense granular flow, which indicates increased drag resistance. This in turn reduces the velocity and the final runout distance. Viscous drag effect is found to reduce the acceleration up to about 30\% in underwater landslides~\citep{Watts2000}. In contrast the loose granular flowing mass experiences less drag, thereby reaching higher kinetic energy and longer runout distances. These mechanisms explain the loose granular masses in fluid spreading almost twice as long as the dense granular masses, as also observed in experiments~\citep{Iverson2000, Schaeffer2008}.

@@

Two-dimensional LB-DEM simulations were performed to understand the effect of initial volume fraction on the behaviour of granular column collapse on slopes submerged in fluid. The cases were run with a granular material with relatively low permeability by adopting varying hydraulic radius. By doing so, it was possible to examine the effect of coupled interaction of pore pressure and interparticle force during different stages of granular column collapse.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[noabbrev,capitalize]{cleveref}
\usepackage{dsfont}
\RequirePackage{amsmath,amsfonts,amssymb}
\RequirePackage{booktabs}

@@

\Abstract{This paper investigates the effect of permeability on the runout characteristics of collapse of  granular columns on slope in fluid. Two-dimensional sub-grain scale numerical simulations are performed to understand the flow dynamics of granular collapse in fluid. The Discrete Element (DEM) technique is coupled with the Lattice Boltzmann Method (LBM), for fluid-grain interactions, to understand the evolution of submerged granular flows. The fluid phase is simulated using Multiple-Relaxation-Time LBM (LBM-MRT) for numerical stability. In order to simulate interconnected pore space in 2D, a reduction in the radius of the grains (hydrodynamic radius) is assumed during LBM computations. The collapse of granular column in fluid is compared with the dry cases to understand the effect of fluid on the runout behaviour. A parametric analysis is performed to assess the influence of the granular characteristics (initial packing) on the evolution of flow and run-out distances for slope angle of \SI{5}{\degree}. In order to understand the effect of permeability on granular flow down a slope angle of \SI{5}{\degree}, the collapse of a granular column with an initial aspect ratio of 0.8 is simulated with different permeabilities. The hydrodynamic radius of a loosely packed granular column is varied from r = 0.7 R (high permeability), 0.75 R, 0.8 R, 0.85 R to 0.9 R (low permeability). The granular flow dynamics is investigated by analysing the effect of hydroplaning, water entrainment and viscous drag on the granular mass. The mechanism of energy dissipation, shape of the flow front, water entrainment and evolution of packing density is used to explain the difference in the flow characteristics of granular column collapse in fluid.}

@@

\section{Granular column collapse on slope in fluid}
\subsection{Problem definition}
In this study, a 2D polydisperse system ($d_{max}/d_{min} = 1.8$) of circular discs in fluid was used to understand the behaviour of granular flows on inclined planes. As shown in~\cref{fig:setup}, an inclined gravity direction was varied to examine the slope angle effect. The soil column was modelled using 1000 discs of density \SI{2650}{\kg\per\cubic\meter} and a contact friction angle of \SI{26}{\degree}. The aspect ratio `a' is defined as the ratio of the initial height ($H_i$) to the width ($L_i$) of the column. A granular column of aspect ratio `a' of 0.8 was used. The collapse of the column was simulated inside a fluid with a density of \SI{1000}{\kg\per\cubic\meter} and a kinematic viscosity of \SI{1e-6}{\square\meter\per\second}. The choice of a 2D geometry has the advantage of cheaper computational effort than a 3D case, making it feasible to simulate very large systems. To model 3D permeability nature of spheres as 2D discs, a reduction in radius: a  hydrodynamic radius r = 0.9R was adopted only for LBM computations, as described earlier. Dry column collapse was also performed to study the effect of hydrodynamic forces on the runout distance. The runout distance is normalised with respect to the initial width of the column. The granular column collapse was allowed to flow down a slope of \SI{5}{\degree}.

\begin{figure}[htpb]
  \centering
\includegraphics[width=0.97\columnwidth]{figs/geometry}
\caption{Underwater granular collapse set-up}
\label{fig:setup}
\end{figure}

@@

\begin{figure}[htpb]
  \centering
\includegraphics[width=0.9\columnwidth]{figs/dense_s5_r09}
\caption{Evolution of a dense sand column (aspect ratio 0.8) on a slope of \SI{5}{\degree}. }
\label{fig:dense-s5-r09}
\end{figure}

@@

\section{Effect of initial density - Loose versus Dense}
\citet{Topin2011} observed from CFD-DEM simulations of granular collapse on a horizontal plane in different viscous fluids that, for a given initial geometry, the run-out distance in the dry case is significantly higher than the submerged case, an observation similar to the experimental results of~\citet{Cassar2005}. For a given geometry, the initial volume fraction of the granular mass has a significant effect on the morphology of the granular deposits in fluid~\citep{Rondon2011a, Pailha2008}. In the dry case, inertia is responsible for the enhanced mobility. In the submerged cases, however, the grain-inertial effects remain negligible because of the predominance of viscous effects on the granular dynamics~\citep{Topin2011}. This could explain the importance of the initial volume fraction on the runout behaviour of submerged granular columns. In this study, we examine the effect of initial packing density on the runout behaviour at \SI{5}{\degree} slope angle in both dry and submerged cases.

@@

\begin{figure}
    \centering

\makebox[\linewidth][c]{
\begin{subfigure}[b]{0.95\textwidth}
	\centering
    \includegraphics[width=0.95\textwidth]{figs/loose_slope/LBM_660_Slope5_r07}
    \caption{r = 0.7 R}
    \label{fig:LBM_660_Slope5_r07}
\end{subfigure}
}\\
\makebox[\linewidth][c]{
\begin{subfigure}[b]{0.95\textwidth}
	\centering
    \includegraphics[width=0.95\textwidth]{figs/loose_slope/LBM_660_Slope5_r075}
    \caption{r = 0.75 R}
    \label{fig:LBM_660_Slope5_r075}
\end{subfigure}
} \\
\makebox[\linewidth][c]{
\begin{subfigure}[b]{0.95\textwidth}
	\centering
    \includegraphics[width=0.95\textwidth]{figs/loose_slope/LBM_660_Slope5_r08}
    \caption{r = 0.8 R}
    \label{fig:LBM_660_Slope5_r08}
\end{subfigure}
}\\
\makebox[\linewidth][c]{
\begin{subfigure}[b]{0.95\textwidth}
	\centering
    \includegraphics[width=0.95\textwidth]{figs/loose_slope/LBM_660_Slope5_r085}
    \caption{r = 0.85 R}
    \label{fig:LBM_660_Slope5_r085}
\end{subfigure}
}\\
\makebox[\linewidth][c]{
\begin{subfigure}[b]{0.95\textwidth}
	\centering
    \includegraphics[width=0.95\textwidth]{figs/loose_slope/LBM_660_Slope5_r09}
    \caption{r = 0.9 R}
    \label{fig:LBM_660_Slope5_r09}
\end{subfigure}
}
\caption{Evolution of the flow front at $t = 3\tau_c$ for different 
permeabilities (loose slope 5\si{\degree}).}
\label{fig:slope_loose_5}
\end{figure}

@@

\subsubsection{Water entrainment}
Comparing the snapshots of the granular flow at time $t=3\tau_c$ for the dense column cases (\cref{fig:dense-s5-r09}) and the loose column cases (\cref{fig:slope_loose_5}), it can be seen that the overall shapes of the flow front are different.~\Cref{fig:slope-outline} shows the boundary of the flowing mass and its evolution with time for the slope angle \si{5}{\degree} case. It shows that the dense granular columns exhibit a more acute angle of attack (flow front) in contrast to a more parabolic profile for the loose granular columns. This difference in angle of flow front changes the ratio of amount of water entering the flowing granular mass to the water interacting with the surface of the granular mass.~\Cref{fig:water_entrainment_loose_dense} shows the horizontal velocity of the fluid and streamlines at the flow front of the dense and loose granular columns for the slope angle of \si{5}{\degree}. The density of streamlines entering the granular front is significantly higher for the loose column case, in comparison to the dense column case. This clearly shows the effect of the shape of flow front on water entrainment. More water entrainment results in increase in pore space (or decrease in solid fraction) at the flow front, which in turn reduces the interparticle shear resistance.

@@

\subsubsection{Turbulence and drag}
Closer observation of the surface of the flow front in the dense case (\cref{fig:dense-s5-r09}) and the loose column (\cref{fig:slope_loose_5}) cases shows that the number of turbulent vortices formed on the surface of the flow is significantly higher in the dense column cases than in the loose column cases. This can be attributed to the triangular profiles of the flowing mass of the dense column cases, in contrast to the more parabolic profiles observed in the loose granular column cases. This increase in the interaction between the flowing granular mass and the neighbouring fluid results in an increased momentum transfer and thus formation of turbulent vortices.~\Cref{fig:water_entrainment_a08_dense} shows higher density of streamlines at the top surface for the dense granular flow, which indicates increased drag resistance. This in turn reduces the velocity and the final runout distance. Viscous drag effect is found to reduce the acceleration up to about 30\% in underwater landslides~\citep{Watts2000}. In contrast the loose granular flowing mass experiences less drag, thereby reaching higher kinetic energy and longer runout distances. These mechanisms explain the loose granular masses in fluid spreading almost twice as long as the dense granular masses, as also observed in experiments~\citep{Iverson2000, Schaeffer2008}.

@@

Two-dimensional LB-DEM simulations were performed to understand the effect of initial volume fraction on the behaviour of granular column collapse on slope of \SI{5}{\degree} submerged in fluid. The cases were run with a granular material with relatively low permeability by adopting varying hydraulic radius. By doing so, it was possible to examine the effect of coupled interaction of pore pressure and interparticle force during different stages of granular column collapse.
